Creating the ``r params$package_name`` R package

### 2025-05-30 create-rhello.Rmd
re-defined the class and operations, right before inprod

```{r}
#' Create a predictor_hybrid object (S3 version)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' along with Jacobians and metadata.
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#' @param jacobian_list A list of Jacobian matrices corresponding to the functional predictors.
#' @param n_basis_list A numeric vector indicating the number of basis functions for each functional predictor.
#' @param n_sample Number of observations (samples).
#' @param n_functional Number of functional predictors.
#' @param n_scalar Number of scalar predictors.
#'
#' @return A list of class \code{predictor_hybrid}.
#' @export
predictor_hybrid <- function(Z, functional_list, jacobian_list,
                             n_basis_list, n_sample, n_functional, n_scalar) {
  stopifnot(is.matrix(Z))
  stopifnot(length(functional_list) == n_functional)
  stopifnot(length(jacobian_list) == n_functional)
  stopifnot(length(n_basis_list) == n_functional)

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      jacobian_list = jacobian_list,
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

Let's write some tests to make sure the function behaves as desired:
```{r}
testthat::test_that("predictor_hybrid constructor works as expected (S3)", {
  Z <- matrix(rnorm(100 * 5), nrow = 100, ncol = 5)
  fd1 <- list(fdobj = "fd1 placeholder")
  fd2 <- list(fdobj = "fd2 placeholder")
  J1 <- matrix(runif(100), nrow = 10)
  J2 <- matrix(runif(200), nrow = 20)

  obj <- predictor_hybrid(
    Z = Z,
    functional_list = list(fd1, fd2),
    jacobian_list = list(J1, J2),
    n_basis_list = c(10, 20),
    n_sample = 100,
    n_functional = 2,
    n_scalar = 5
  )

  testthat::expect_s3_class(obj, "predictor_hybrid")
  testthat::expect_equal(nrow(obj$Z), 100)
  testthat::expect_equal(ncol(obj$Z), 5)
  testthat::expect_equal(length(obj$functional_list), 2)
  testthat::expect_equal(length(obj$jacobian_list), 2)
  testthat::expect_equal(obj$n_basis_list, c(10, 20))
  testthat::expect_equal(obj$n_sample, 100)
  testthat::expect_equal(obj$n_functional, 2)
  testthat::expect_equal(obj$n_scalar, 5)
})

@@

@@

```{r}
#' Add two predictor_hybrid objects
#'
#' Performs element-wise addition of two predictor_hybrid objects.
#' Functional predictors are combined using `plus.fd()` and `times.fd()`.
#'
#' @param input A predictor_hybrid object.
#' @param other Another predictor_hybrid object.
#' @param alpha A numeric scalar multiplier for `other`.
#'
#' @return A new predictor_hybrid object representing the sum.
#' @export
add.predictor_hybrid <- function(input, other, alpha = 1) {
  if (input$n_functional != other$n_functional || input$n_scalar != other$n_scalar) {
    stop("Mismatch in number of predictors.")
  }

  # Optionally check that functional bases are the same
  # if (!is_same_basis(input, other)) stop("Different functional bases.")

  new_functional_list <- vector("list", input$n_functional)
  for (i in seq_len(input$n_functional)) {
    new_functional_list[[i]] <- plus.fd(
      input$functional_list[[i]],
      times.fd(alpha, other$functional_list[[i]])
    )
  }

  new_Z <- input$Z + alpha * other$Z

  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list,
    jacobian_list = input$jacobian_list,  # Assumes identical structure
    n_basis_list = input$n_basis_list,
    n_sample = input$n_sample,
    n_functional = input$n_functional,
    n_scalar = input$n_scalar
  )
}

```{r}
testthat::test_that("add.predictor_hybrid works for matching predictors", {
  suppressPackageStartupMessages(library(fda))

  Z1 <- matrix(1, nrow = 3, ncol = 2)
  Z2 <- matrix(2, nrow = 3, ncol = 2)

  basis <- fda::create.bspline.basis(c(0, 1), 5)
  fd1 <- fda::fd(coef = matrix(1, 5, 3), basisobj = basis)
  fd2 <- fda::fd(coef = matrix(2, 5, 3), basisobj = basis)

  obj1 <- predictor_hybrid(Z1, list(fd1), list(matrix(0)), c(5), 3, 1, 2)
  obj2 <- predictor_hybrid(Z2, list(fd2), list(matrix(0)), c(5), 3, 1, 2)

  obj3 <- add.predictor_hybrid(obj1, obj2, alpha = 1)

  testthat::expect_s3_class(obj3, "predictor_hybrid")
  testthat::expect_equal(obj3$Z, matrix(3, nrow = 3, ncol = 2))

  expected_coef <- matrix(3, 5, 3)
  testthat::expect_equal(
    unname(coef(obj3$functional_list[[1]])),
    unname(expected_coef)
  )
})

@@

```{r}
#' Multiply a predictor_hybrid object by a scalar
#'
#' Performs scalar multiplication of both scalar and functional components 
#' of a `predictor_hybrid` object. Functional predictors are scaled using 
#' `times.fd()` from the `fda` package.
#'
#' @param input A `predictor_hybrid` object.
#' @param scalar A numeric value to multiply all components by.
#'
#' @return A new `predictor_hybrid` object scaled by `scalar`.
#' @export
scalar_mul.predictor_hybrid <- function(input, scalar) {
  new_functional_list <- lapply(
    input$functional_list,
    function(fd_obj) times.fd(scalar, fd_obj)
  )

  new_Z <- scalar * input$Z

  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list,
    jacobian_list = input$jacobian_list,
    n_basis_list = input$n_basis_list,
    n_sample = input$n_sample,
    n_functional = input$n_functional,
    n_scalar = input$n_scalar
  )
}
```

@@

  # Scalar matrix
  Z <- matrix(2, nrow = 3, ncol = 2)

  # Create a shared basis for both functional predictors
  basis <- create.bspline.basis(c(0, 1), 5)

  # Two identical fd objects
  fd1 <- fd(coef = matrix(2, 5, 3), basisobj = basis)
  fd2 <- fd(coef = matrix(2, 5, 3), basisobj = basis)

  # Construct predictor_hybrid object
  obj <- predictor_hybrid(
    Z = Z,
    functional_list = list(fd1, fd2),
    jacobian_list = list(matrix(0), matrix(0)),
    n_basis_list = c(5, 5),
    n_sample = 3,
    n_functional = 2,
    n_scalar = 2
  )

  # Perform scalar multiplication
  scaled <- scalar_mul.predictor_hybrid(obj, 4)

  testthat::expect_s3_class(scaled, "predictor_hybrid")

  # Scalar part should be scaled
  testthat::expect_equal(scaled$Z, Z * 4)

  # Each functional part should be scaled
  expected_coef <- matrix(8, 5, 3)  # 2 * 4
  for (fd_scaled in scaled$functional_list) {
    testthat::expect_equal(
      unname(coef(fd_scaled)),
      unname(expected_coef)
    )
  }
})
```
### inprod.predictor_hybrid
The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)}) $ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as $\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k$ with norm $\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}$.

@@

>>>>

```{r}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

  n_sample <- nrow(Z)
  n_scalar <- ncol(Z)
  n_functional <- length(functional_list)

  gram_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

  for (i in seq_len(n_functional)) {
    fd_i <- functional_list[[i]]
    stopifnot(inherits(fd_i, "fd"))

    n_fd_sample <- ncol(coef(fd_i))
    if (n_fd_sample != n_sample) {
      stop(sprintf("Functional predictor %d has %d replicates, but Z has %d rows.",
                   i, n_fd_sample, n_sample))
    }

    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    n_basis_list[i] <- basis_i$nbasis
  }

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      gram_list = gram_list,
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}

@@

Let's write some tests to make sure the function behaves as desired:
```{r}
testthat::test_that("predictor_hybrid constructor works as expected (S3, updated)", {
  library(fda)

  # Create scalar matrix
  Z <- matrix(rnorm(100 * 5), nrow = 100, ncol = 5)

  # Create two fd objects with 100 replications
  basis1 <- create.bspline.basis(c(0, 1), nbasis = 10)
  basis2 <- create.bspline.basis(c(0, 1), nbasis = 20)

  coef1 <- matrix(rnorm(10 * 100), nrow = 10)  # 10 basis × 100 samples
  coef2 <- matrix(rnorm(20 * 100), nrow = 20)  # 20 basis × 100 samples

  fd1 <- fd(coef = coef1, basisobj = basis1)
  fd2 <- fd(coef = coef2, basisobj = basis2)

  # Construct object
  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Assertions
  testthat::expect_s3_class(obj, "predictor_hybrid")
  testthat::expect_equal(nrow(obj$Z), 100)
  testthat::expect_equal(ncol(obj$Z), 5)
  testthat::expect_equal(length(obj$functional_list), 2)
  testthat::expect_equal(obj$n_basis_list, c(10, 20))
  testthat::expect_equal(obj$n_sample, 100)
  testthat::expect_equal(obj$n_functional, 2)
  testthat::expect_equal(obj$n_scalar, 5)
  testthat::expect_equal(dim(obj$gram_list[[1]]), c(10, 10))
  testthat::expect_equal(dim(obj$gram_list[[2]]), c(20, 20))
})
```

@@

```{r}
add.predictor_hybrid <- function(input, other, alpha = 1) {
  if (!inherits(input, "predictor_hybrid") || !inherits(other, "predictor_hybrid")) {
    stop("Both arguments must be of class 'predictor_hybrid'.")
  }

  if (input$n_functional != other$n_functional || input$n_scalar != other$n_scalar) {
    stop("Mismatch in number of functional or scalar predictors.")
  }

  # Check broadcasting compatibility
  n1 <- input$n_sample
  n2 <- other$n_sample
  if (n1 != n2 && !(n1 == 1 || n2 == 1)) {
    stop("Sample sizes are incompatible for broadcasting.")
  }

  # Safe access to is.eqbasis()
  is_eqbasis <- getFromNamespace("is.eqbasis", "fda")

  for (i in seq_len(input$n_functional)) {
    if (!is_eqbasis(input$functional_list[[i]]$basis, other$functional_list[[i]]$basis)) {
      stop("Functional predictors must have the same basis.")
    }
  }

  # Broadcast if needed
  broadcast <- function(obj, n_target) {
    if (obj$n_sample == n_target) return(obj)
    obj$Z <- matrix(rep(obj$Z, each = n_target), nrow = n_target, byrow = TRUE)
    obj$functional_list <- lapply(obj$functional_list, function(fd_obj) {
      fd(coef = matrix(rep(coef(fd_obj), n_target), nrow = nrow(coef(fd_obj))),
         basisobj = fd_obj$basis)
    })
    obj$n_sample <- n_target
    obj
  }

  # Determine output sample size
  n_out <- max(n1, n2)
  input <- broadcast(input, n_out)
  other <- broadcast(other, n_out)

  # Combine functional predictors
  new_functional_list <- Map(
    function(f1, f2) plus.fd(f1, times.fd(alpha, f2)),
    input$functional_list,
    other$functional_list
  )

  # Combine scalar predictors
  new_Z <- input$Z + alpha * other$Z

  # Construct new object
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
}
```

Here are the unit tests:
| Case                    | Description                         | Expected Outcome |
| ----------------------- | ----------------------------------- | ---------------- |
| ✅ Same sample size      | Functional and scalar values summed | Pass             |
| ✅ Broadcast one → many  | One-sample input broadcasted        | Pass             |
| ✅ Broadcast many ← one  | One-sample other broadcasted        | Pass             |
| ❌ Basis mismatch        | Different functional basis          | Error            |
| ❌ Irregular sample size | Neither matches nor is singleton    | Error            |
| ❌ Different dimensions  | Functional or scalar count mismatch | Error            |
```{r}
testthat::test_that("add.predictor_hybrid handles broadcasting and basis compatibility", {
  suppressPackageStartupMessages(library(fda))

  # Create shared B-spline basis
  basis <- create.bspline.basis(c(0, 1), nbasis = 5)

  # Helper to generate a valid coefficient matrix
  coef_mat <- function(val, n_sample, basis_obj) {
    nbasis <- basis_obj$nbasis
    matrix(val, nrow = nbasis, ncol = n_sample)
  }

  # Helper to create a list of fd objects (2 functional predictors)
  fd_list <- function(val, n_sample, basis_obj = basis) {
    list(
      fd(coef = coef_mat(val, n_sample, basis_obj), basisobj = basis_obj),
      fd(coef = coef_mat(val, n_sample, basis_obj), basisobj = basis_obj)
    )
  }

  # Scalar predictors
  Z1 <- matrix(1, nrow = 1, ncol = 2)
  Z3 <- matrix(3, nrow = 3, ncol = 2)

  # Functional predictors (same basis)
  F1 <- fd_list(1, 1)
  F3 <- fd_list(3, 3)

  # Create predictor_hybrid objects
  x1 <- predictor_hybrid(Z1, F1)  # 1 sample
  x3 <- predictor_hybrid(Z3, F3)  # 3 samples

  # Broadcast x1 to match x3
  result <- add.predictor_hybrid(x1, x3, alpha = 1)

  # Check metadata
  testthat::expect_s3_class(result, "predictor_hybrid")
  testthat::expect_equal(result$n_sample, 3)
  testthat::expect_equal(result$n_scalar, 2)
  testthat::expect_equal(result$n_functional, 2)

  # Check scalar component
  expected_Z <- matrix(4, nrow = 3, ncol = 2)  # 1 + 3 = 4
  testthat::expect_equal(result$Z, expected_Z)

  # Check functional component
  for (fd_out in result$functional_list) {
    testthat::expect_equal(unname(coef(fd_out)), matrix(4, nrow = 5, ncol = 3))
  }

  # Test symmetry: broadcasting x3 + x1
  result2 <- add.predictor_hybrid(x3, x1, alpha = 1)
  testthat::expect_equal(result2$Z, expected_Z)
  for (fd_out in result2$functional_list) {
    testthat::expect_equal(unname(coef(fd_out)), matrix(4, nrow = 5, ncol = 3))
  }

  # Incompatible sample sizes (should fail)
  x_bad <- predictor_hybrid(matrix(1, 2, 2), fd_list(1, 2))
  testthat::expect_error(add.predictor_hybrid(x3, x_bad), "incompatible for broadcasting")

  # Incompatible functional basis (should fail)
  bad_basis <- create.bspline.basis(c(0, 1), nbasis = 6)
  bad_fd <- fd_list(2, 3, bad_basis)
  x_basis_mismatch <- predictor_hybrid(Z3, bad_fd)
  testthat::expect_error(add.predictor_hybrid(x3, x_basis_mismatch), "same basis")
})

@@

```{r}
#' Multiply a predictor_hybrid object by a scalar
#'
#' Performs scalar multiplication on both the scalar and functional components
#' of a `predictor_hybrid` object. Functional predictors are scaled using 
#' `times.fd()` from the `fda` package.
#'
#' @param input A `predictor_hybrid` object.
#' @param scalar A numeric value to multiply all components by.
#'
#' @return A new `predictor_hybrid` object scaled by `scalar`.
#' @export
scalar_mul.predictor_hybrid <- function(input, scalar) {
  if (!inherits(input, "predictor_hybrid")) {
    stop("Input must be of class 'predictor_hybrid'.")
  }
  if (!is.numeric(scalar) || length(scalar) != 1) {
    stop("Scalar must be a single numeric value.")
  }

  # Scale functional components
  new_functional_list <- lapply(input$functional_list, function(fd_obj) {
    times.fd(scalar, fd_obj)
  })

  # Scale scalar predictors
  new_Z <- scalar * input$Z

  # Reconstruct hybrid object
  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list
  )
}
```

@@

  # Scalar part (3 observations, 2 scalar predictors)
  Z <- matrix(2, nrow = 3, ncol = 2)

  # Functional part: shared B-spline basis
  basis <- create.bspline.basis(c(0, 1), nbasis = 5)
  fd1 <- fd(coef = matrix(2, nrow = 5, ncol = 3), basisobj = basis)
  fd2 <- fd(coef = matrix(2, nrow = 5, ncol = 3), basisobj = basis)

  # Construct hybrid object using new simplified constructor
  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Apply scalar multiplication
  scaled <- scalar_mul.predictor_hybrid(obj, scalar = 4)

  # Check class
  testthat::expect_s3_class(scaled, "predictor_hybrid")

  # Check scalar predictor was scaled
  testthat::expect_equal(scaled$Z, Z * 4)

  # Check each functional component's coefficients
  expected_coef <- matrix(8, nrow = 5, ncol = 3)
  for (fd_obj in scaled$functional_list) {
    testthat::expect_equal(
      unname(coef(fd_obj)),
      unname(expected_coef)
    )
  }
})
```
### inprod.predictor_hybrid
The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)}) $ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as $\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k$ with norm $\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}$.

@@

### compute_gram_matrix
```{r}
#' Compute the Gram matrix for a basis object
#'
#' @param basis A basis object of class `basisfd` from the `fda` package
#' @return A square matrix where entry (i,j) is the inner product of basis functions i and j
#' @export
compute_gram_matrix <- function(basis) {
  stopifnot(inherits(basis, "basisfd"))
  inprod(basis, basis)  # computes ∫ b_i(t) b_j(t) dt
}

We check the following cases:
| Case | Test                 | Expectation                             |
| ---- | -------------------- | --------------------------------------- |
| 1    | Orthogonal (Fourier) | Gram matrix ≈ diagonal                  |
| 2    | Constant             | Single value equals interval length (1) |
| 3    | Invalid              | Raises informative error                |
| 4    | Narrow support       | Mostly low off-diagonal entries         |

```{r}
testthat::test_that("Gram matrix edge cases", {
  library(fda)

  # Case 1: Fourier basis (orthogonal over [0, 1])
  fb <- create.fourier.basis(rangeval = c(0, 1), nbasis = 5)
  J_fourier <- compute_gram_matrix(fb)
  testthat::expect_equal(J_fourier, diag(diag(J_fourier)), tolerance = 1e-10)  # Nearly diagonal

  # Case 2: Constant basis (1 basis function)
  cb <- create.constant.basis(rangeval = c(0, 1))
  J_const <- compute_gram_matrix(cb)
  testthat::expect_equal(dim(J_const), c(1, 1))
  testthat::expect_equal(J_const[1, 1], 1)

  # Case 3: Invalid input
  testthat::expect_error(compute_gram_matrix("not a basisfd"))

  # Case 4: Highly localized B-splines
  # Simulate nearly disjoint support (narrow B-splines)
  narrow_bs <- create.bspline.basis(rangeval = c(0, 1), nbasis = 10, norder = 2)
  J_narrow <- compute_gram_matrix(narrow_bs)
  testthat::expect_true(all(abs(J_narrow[upper.tri(J_narrow)]) < 0.5))  # off-diagonals small
})
